---
title: "Supplementary Materials"

format:
  apaquarto-pdf:
    toc: true
    toc-depth: 2

author:
  - name: Matteo Manente
    affiliations:
      - name: University of Padova
        id: 1
        department: Department of Developmental Psychology and Socialization
        address: Via Venezia 8
        city: Padova
        state: Italy
        postal-code: 35131

  - name: Livio Finos
    affiliations:
      - name: University of Padova
        id: 2
        department: Department of Statistical Sciences
        address: Via Battisti 241
        city: Padova
        state: Italy
        postal-code: 35121

  - name: Gianmarco Altoè
    affiliations:
      - ref: 1

  - name: Filippo Gambarota
    affiliations:
      - ref: 1
    corresponding: true
    email: filippo.gambarota@unipd.it

bibliography:
  - "pimma.bib"
  - "references.bib"

floatsintext: true
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(metafor)
library(here)
library(ggh4x)
library(flipmeta)

knitr::opts_chunk$set(
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  fig.align = "center",
  dev = "cairo_pdf"
)

mtheme <- function() {
  theme_minimal(16) +
    theme(
      legend.position = "bottom"
    )
}

theme_set(mtheme())

# -----------------------------------------------------------------------
# Reproducibility information
# -----------------------------------------------------------------------

operating_system <- paste(
  Sys.info()[["sysname"]],
  Sys.info()[["release"]],
  Sys.info()[["machine"]]
)

git_commit <- tryCatch(
  system2(
    "git",
    c("rev-parse", "HEAD"),
    stdout = TRUE
  ),
  error = function(e) NA_character_
)

quarto_version <- tryCatch(
  system2(
    "quarto",
    "--version",
    stdout = TRUE
  ),
  error = function(e) NA_character_
)

software_versions <- tibble(
  Software = c(
    "R",
    "Quarto",
    "metafor",
    "flipmeta",
    "flipscores",
    "dplyr",
    "tidyr",
    "ggplot2"
  ),
  Version = c(
    paste0(
      R.version$major,
      ".",
      R.version$minor
    ),
    quarto_version,
    as.character(packageVersion("metafor")),
    as.character(packageVersion("flipmeta")),
    as.character(packageVersion("flipscores")),
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("tidyr")),
    as.character(packageVersion("ggplot2"))
  )
)

# -----------------------------------------------------------------------
# Simulation results
# -----------------------------------------------------------------------

sim <- readRDS(
  here::here(
    "simulation",
    "res-clean.rds"
  )
)

remove_redundant <- function(data, cols = NULL) {

  cols <- if (is.null(cols)) names(data) else cols

  keep <- rep(
    TRUE,
    ncol(data)
  )

  names(keep) <- names(data)

  keep[cols] <- sapply(
    data[cols],
    function(x) length(unique(x)) > 1
  )

  data[, keep, drop = FALSE]
}

sim_long <- sim |>
  remove_redundant() |>
  pivot_longer(
    starts_with("pval"),
    names_to = "inf",
    values_to = "pval"
  )

inf_lbl <- c(
  "pval_adhoc" = "mKH",
  "pval_knha" = "KH",
  "pval_rob" = "RVE",
  "pval_flip" = "flipmeta",
  "pval_wald" = "Wald"
)

# Type I error by individual simulation condition

type1_supp <- sim_long |>
  filter(
    bX == 0
  ) |>
  group_by(
    rXZ,
    tau2,
    zeta,
    k,
    n,
    method,
    inf
  ) |>
  summarise(
    rejection_rate = mean(
      pval <= 0.05,
      na.rm = TRUE
    ),
    nsim = n(),
    .groups = "drop"
  ) |>
  mutate(
    method = ifelse(
      method == "EE",
      "EE",
      "RE--REML"
    ),
    inf = inf_lbl[inf]
  )

# Identify method-condition combinations with adequate Type I error control

type1_ok <- type1_supp |>
  mutate(
    type1_upper = 0.05 +
      qnorm(0.975) *
      sqrt(
        0.05 * 0.95 / nsim
      ),
    type1_ok = rejection_rate <= type1_upper
  ) |>
  select(
    rXZ,
    tau2,
    zeta,
    k,
    n,
    method,
    inf,
    type1_ok
  )

# Power by individual simulation condition

power_supp <- sim_long |>
  filter(
    bX != 0
  ) |>
  group_by(
    rXZ,
    tau2,
    zeta,
    k,
    n,
    method,
    inf
  ) |>
  summarise(
    power = mean(
      pval <= 0.05,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  mutate(
    method = ifelse(
      method == "EE",
      "EE",
      "RE--REML"
    ),
    inf = inf_lbl[inf]
  ) |>
  left_join(
    type1_ok,
    by = c(
      "rXZ",
      "tau2",
      "zeta",
      "k",
      "n",
      "method",
      "inf"
    )
  ) |>
  filter(
    type1_ok
  )

# -----------------------------------------------------------------------
# CBT multiverse
# -----------------------------------------------------------------------

multi <- readRDS(
  here::here(
    "application",
    "cbt-depression",
    "results",
    "cbt-dep-multi.rds"
  )
)
```

# S1. Reproducibility and software

All analyses were conducted in R. Meta-analytic models were fitted using `metafor`, and score-based sign-flipping inference was performed using `flipmeta`. Random-number seeds were fixed for all simulation and resampling procedures.

The source code used to generate the simulations, empirical application, figures, and tables is available in the accompanying repository. The manuscript and supplementary materials were generated directly from Quarto source files.

```{r}
#| label: tbl-software
#| tbl-cap: "Main software versions used to reproduce the analyses."

software_versions |>
  knitr::kable(
    align = c(
      "l",
      "l"
    )
  )
```

The supplementary materials were rendered on `r operating_system` from Git commit `r git_commit`.

# S2. Computational comparison with permutation testing

The main manuscript notes that score sign flipping avoids repeated refitting of the complete meta-regression model. We illustrate the practical computational difference using a mixed-effects meta-regression with 100 independent studies and three predictors.

Both procedures used 1,000 resampling transformations. The benchmark is intended only as an illustration; computation times depend on the model, hardware, software versions, and number of resampling transformations.

```{r}
#| label: computational-benchmark
#| cache: true

sim_meta <- function(
  k,
  mu,
  tau2,
  n1,
  n2 = NULL,
  v = 1
) {

  if (is.null(n2)) {
    n2 <- n1
  }

  deltai <- rnorm(
    k,
    0,
    sqrt(tau2)
  )

  yi <- mu +
    deltai +
    rnorm(
      k,
      0,
      sqrt(
        v * (
          1 / n1 +
            1 / n2
        )
      )
    )

  vi <- (
    rchisq(
      k,
      n1 + n2 - 2
    ) /
      (
        n1 +
          n2 -
          2
      )
  ) *
    v *
    (
      1 / n1 +
        1 / n2
    )

  data.frame(
    id = seq_len(k),
    yi,
    vi
  )
}

elapsed <- function(expr) {

  start <- Sys.time()

  res <- eval(
    substitute(expr),
    envir = parent.frame()
  )

  time <- as.numeric(
    difftime(
      Sys.time(),
      start,
      units = "secs"
    )
  )

  list(
    result = res,
    time = time
  )
}

set.seed(2026)

k_benchmark <- 100
B_benchmark <- 1000

x1 <- rnorm(k_benchmark)
x2 <- rnorm(k_benchmark)
x3 <- rnorm(k_benchmark)

dat_benchmark <- sim_meta(
  k = k_benchmark,
  mu = 0.1 +
    0.5 * x1 +
    0.1 * x2 +
    0.4 * x3,
  tau2 = 0.1,
  n1 = 50
)

dat_benchmark <- cbind(
  dat_benchmark,
  x1,
  x2,
  x3
)

fit_benchmark <- metafor::rma(
  yi,
  vi,
  data = dat_benchmark,
  mods = ~ x1 + x2 + x3
)

set.seed(2026)

perm_rma <- elapsed(
  metafor::permutest(
    fit_benchmark,
    exact = FALSE,
    iter = B_benchmark,
    progbar = FALSE
  )
)

set.seed(2026)

flip_rma <- elapsed(
  flipmeta::flipmeta(
    fit_benchmark,
    B = B_benchmark,
    progress = FALSE
  )
)

benchmark_results <- tibble(
  Procedure = c(
    "metafor::permutest()",
    "flipmeta"
  ),
  `Time (seconds)` = c(
    perm_rma$time,
    flip_rma$time
  )
)
```

```{r}
#| label: tbl-computational-benchmark
#| tbl-cap: "Illustrative computation times for 1,000 resampling transformations."

benchmark_results |>
  mutate(
    `Time (seconds)` = sprintf(
      "%.2f",
      `Time (seconds)`
    )
  ) |>
  knitr::kable(
    align = c(
      "l",
      "c"
    )
  )
```

In this example, `metafor::permutest()` required approximately `r round(perm_rma$time, 1)` seconds, compared with `r round(flip_rma$time, 2)` seconds for `flipmeta`.

# S3. Additional simulation results

The main manuscript reports simulation results averaged over study sample size and moderator correlation. Here we show the corresponding results separately across these conditions.

## S3.1 Type I error

```{r}
#| label: fig-type1-detailed
#| fig-cap: "Empirical Type I error rates across study sample sizes and moderator correlations. The horizontal dotted line indicates the nominal significance level of .05. Wald is omitted for readability because it showed substantial Type I error inflation in several conditions."
#| fig-width: 11
#| fig-height: 9

type1_supp |>
  filter(
    inf != "Wald"
  ) |>
  mutate(
    rho_label = paste0(
      "rho[XZ] = ",
      rXZ
    ),
    n_label = paste0(
      "n = ",
      n
    ),
    tau_label = paste0(
      "tau0^2 = ",
      tau2,
      ", zeta = ",
      round(zeta, 2)
    )
  ) |>
  ggplot(
    aes(
      x = k,
      y = rejection_rate,
      color = inf,
      linetype = inf
    )
  ) +
  geom_hline(
    yintercept = 0.05,
    linetype = "dotted",
    linewidth = 0.5
  ) +
  geom_line(
    linewidth = 0.7
  ) +
  geom_point(
    size = 1.5
  ) +
  facet_grid(
    tau_label + rho_label ~ method + n_label
  ) +
  labs(
    x = "Number of studies (k)",
    y = "Type I error",
    color = NULL,
    linetype = NULL
  ) +
  theme_bw(
    base_size = 10
  ) +
  theme(
    legend.position = "bottom"
  )
```

The disaggregated results show the conditions underlying the averages reported in the main manuscript and allow calibration to be inspected across study sample size and moderator correlation.

## S3.2 Statistical power

As in the main manuscript, power is shown only for method-condition combinations that adequately controlled Type I error under the corresponding null condition.

```{r}
#| label: fig-power-detailed
#| fig-cap: "Empirical power across study sample sizes and moderator correlations. Results are shown only for method-condition combinations that adequately controlled Type I error under the corresponding null condition."
#| fig-width: 11
#| fig-height: 9

power_supp |>
  filter(
    inf != "Wald"
  ) |>
  mutate(
    rho_label = paste0(
      "rho[XZ] = ",
      rXZ
    ),
    n_label = paste0(
      "n = ",
      n
    ),
    tau_label = paste0(
      "tau0^2 = ",
      tau2,
      ", zeta = ",
      round(zeta, 2)
    )
  ) |>
  ggplot(
    aes(
      x = k,
      y = power,
      color = inf,
      linetype = inf
    )
  ) +
  geom_line(
    linewidth = 0.7
  ) +
  geom_point(
    size = 1.5
  ) +
  facet_grid(
    tau_label + rho_label ~ method + n_label
  ) +
  labs(
    x = "Number of studies (k)",
    y = "Power",
    color = NULL,
    linetype = NULL
  ) +
  theme_bw(
    base_size = 10
  ) +
  theme(
    legend.position = "bottom"
  )
```

These results complement the aggregated power results reported in the main manuscript.


# References

::: {#refs}

:::
