---
title: "Post-selection Inference in Multiverse Meta-Analysis: The PIMMA Framework"
subtitle: "Supplementary Materials"
format: 
  html:
    page-layout: full
    grid:
      body-width: 1400px
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(ggplot2)
library(metafor)
library(here)
library(ggh4x)
library(flipmeta)
library(patchwork)
library(reactable)
library(scales)
library(htmltools)

knitr::opts_chunk$set(
  warning = FALSE,
  echo = FALSE,
  message = FALSE,
  fig.align = "center",
  dev = "svg"
)

mtheme <- function() {
  theme_minimal(16) +
    theme(
      legend.position = "bottom"
    )
}

theme_set(mtheme())

inf_lbl <- c(
  "pval_adhoc" = "mKH",
  "pval_knha"  = "KH",
  "pval_rob"   = "RVE",
  "pval_flip"  = "flipmeta",
  "pval_wald"  = "Wald"
)

sim <- readRDS(
  here::here(
    "simulation",
    "res-clean.rds"
  )
)

remove_redundant <- function(data, cols = NULL) {

  cols <- if (is.null(cols)) names(data) else cols

  keep <- rep(
    TRUE,
    ncol(data)
  )

  names(keep) <- names(data)

  keep[cols] <- sapply(
    data[cols],
    function(x) length(unique(x)) > 1
  )

  data[, keep, drop = FALSE]
}

multi <- readRDS(
  here::here(
    "application",
    "cbt-depression",
    "results",
    "cbt-dep-multi.rds"
  )
)

res <- multi$res

make_table <- function(data, id, palette, threshold = NULL, domain = c(0, 1)) {
  
  tab <- data |>
    remove_redundant() |>
    pivot_wider(
      names_from = name,
      values_from = p
    ) |>
    mutate(
      across(where(is.numeric), \(x) round(x, 3))
    )
  
  result_cols <- intersect(
    unname(inf_lbl),
    names(tab)
  )
  
  dropdown_cols <- setdiff(names(tab), result_cols)
  
  pal <- col_numeric(
    palette = palette,
    domain = domain
  )
  
  cell_color <- function(value) {
    if (!is.null(threshold) && value <= threshold) {
      "white"
    } else {
      pal(pmin(value, max(domain)))
    }
  }
  
  columns <- setNames(
    lapply(
      names(tab),
      \(x) {
        
        is_num <- is.numeric(tab[[x]])
        is_int <- is_num &&
          all(is.na(tab[[x]]) | tab[[x]] %% 1 == 0)
        
        colDef(
          format = if (is_num) {
            colFormat(
              digits = if (x %in% result_cols) 3 else if (is_int) 0 else 3
            )
          },
          
          style = if (x %in% result_cols) {
            function(value) {
              list(background = cell_color(value))
            }
          },
          
          filterInput = if (x %in% dropdown_cols) {
            function(values, name) {
              tags$select(
                onchange = sprintf(
                  "Reactable.setFilter('%s', '%s', this.value)",
                  id,
                  name
                ),
                tags$option("All", value = ""),
                lapply(
                  sort(unique(values)),
                  \(v) tags$option(v, value = v)
                )
              )
            }
          }
        )
      }
    ),
    names(tab)
  )
  
  reactable(
    tab,
    filterable = TRUE,
    searchable = TRUE,
    sortable = TRUE,
    pagination = FALSE,
    columns = columns,
    elementId = id
  )
}
```

# S1. Reproducibility and software

All analyses were conducted in R. Meta-analytic models were fitted using `metafor`, and score-based sign-flipping inference was performed using `flipmeta`. Random-number seeds were fixed for all simulation and resampling procedures.

The source code used to generate the simulations, empirical application, figures, and tables is available in the accompanying repository. The manuscript and supplementary materials were generated directly from Quarto source files.

# S2. Computational comparison with permutation testing

The main manuscript notes that score sign flipping avoids repeated refitting of the complete meta-regression model. We illustrate the practical computational difference using a mixed-effects meta-regression with 100 independent studies and three predictors.

Both procedures used 1,000 resampling transformations. The benchmark is intended only as an illustration; computation times depend on the model, hardware, software versions, and number of resampling transformations. The timing results are summarized in @tbl-computational-benchmark.

```{r}
#| label: computational-benchmark
#| cache: true

sim_meta <- function(
  k,
  mu,
  tau2,
  n1,
  n2 = NULL,
  v = 1
) {

  if (is.null(n2)) {
    n2 <- n1
  }

  deltai <- rnorm(
    k,
    0,
    sqrt(tau2)
  )

  yi <- mu +
    deltai +
    rnorm(
      k,
      0,
      sqrt(
        v * (
          1 / n1 +
            1 / n2
        )
      )
    )

  vi <- (
    rchisq(
      k,
      n1 + n2 - 2
    ) /
      (
        n1 +
          n2 -
          2
      )
  ) *
    v *
    (
      1 / n1 +
        1 / n2
    )

  data.frame(
    id = seq_len(k),
    yi,
    vi
  )
}

elapsed <- function(expr) {

  start <- Sys.time()

  res <- eval(
    substitute(expr),
    envir = parent.frame()
  )

  time <- as.numeric(
    difftime(
      Sys.time(),
      start,
      units = "secs"
    )
  )

  list(
    result = res,
    time = time
  )
}

set.seed(2026)

k_benchmark <- 100
B_benchmark <- 1000

x1 <- rnorm(k_benchmark)
x2 <- rnorm(k_benchmark)
x3 <- rnorm(k_benchmark)

dat_benchmark <- sim_meta(
  k = k_benchmark,
  mu = 0.1 +
    0.5 * x1 +
    0.1 * x2 +
    0.4 * x3,
  tau2 = 0.1,
  n1 = 50
)

dat_benchmark <- cbind(
  dat_benchmark,
  x1,
  x2,
  x3
)

fit_benchmark <- metafor::rma(
  yi,
  vi,
  data = dat_benchmark,
  mods = ~ x1 + x2 + x3
)

set.seed(2026)

perm_rma <- elapsed(
  metafor::permutest(
    fit_benchmark,
    exact = FALSE,
    iter = B_benchmark,
    progbar = FALSE
  )
)

set.seed(2026)

flip_rma <- elapsed(
  flipmeta::flipmeta(
    fit_benchmark,
    B = B_benchmark,
    progress = FALSE
  )
)

benchmark_results <- tibble(
  Procedure = c(
    "metafor::permutest()",
    "flipmeta"
  ),
  `Time (seconds)` = c(
    perm_rma$time,
    flip_rma$time
  )
)
```

```{r}
#| label: tbl-computational-benchmark
#| tbl-cap: "Illustrative computation times for 1,000 resampling transformations."

benchmark_results |>
  mutate(
    `Time (seconds)` = sprintf(
      "%.2f",
      `Time (seconds)`
    )
  ) |>
  knitr::kable(
    align = c(
      "l",
      "c"
    )
  )
```

In this example, `metafor::permutest()` required approximately `r round(perm_rma$time, 1)` seconds, compared with `r round(flip_rma$time, 2)` seconds for `flipmeta` (see @tbl-computational-benchmark).

# S3. Additional simulation results

This section presents full condition-level empirical results from the simulation study evaluating specification-level inference. The interactive tables allow filtering across simulation conditions, including the number of studies ($k$), study sample size ($n$), baseline heterogeneity ($\tau^2$), unobserved confounder correlation ($\zeta$), moderator correlation ($r_{XZ}$), and regression coefficients ($\beta_0, \beta_X, \beta_Z$).

Five inferential procedures are evaluated:
- **mKH**: Modified Knapp--Hartung adjustment
- **KH**: Knapp--Hartung adjustment
- **RVE**: Robust Variance Estimation with CR2 adjustment and Satterthwaite degrees of freedom
- **flipmeta**: Score-based sign-flipping inference
- **Wald**: Standard Wald $z$-test

```{r}
#| echo: false
#| warning: false
#| message: false

sim_res <- sim |> 
  pivot_longer(starts_with("pval")) |> 
  group_by(nX, nZ, b0, bX, bZ, rXZ, tau2, zeta, k, n, method, name) |> 
  summarise(
    p = mean(value <= 0.05),
    .groups = "drop"
  ) |>
  mutate(
    cond = if_else(bX == 0, "type1", "power"),
    name = inf_lbl[name]
  )

power <- filter(sim_res, cond == "power")
type1 <- filter(sim_res, cond == "type1")
```

## Power

Empirical power under $\beta_X = 0.2$ across simulated conditions. Green shading highlights higher power values (scale from 0 to 1).

```{r}
#| echo: false

make_table(
  power,
  id = "power-table",
  palette = c("white", "#63BE7B"),
  domain = c(0, 1)
)
```

## Type I error

Empirical Type I error rates under $\beta_X = 0$ across simulated conditions at nominal $\alpha = 0.05$. Red shading highlights inflated Type I error rates exceeding 0.05.

```{r}
#| echo: false

make_table(
  type1,
  id = "type1-table",
  palette = c("white", "#F8696B"),
  threshold = 0.05,
  domain = c(0.05, 0.3)
)
```

# S4. CBT depression specification curve

The specification curve (@fig-specification-curve) displays the complete set of unique analyses, ordered by their pooled effect estimate. “Non-discriminating” indicates that multiple nominal levels of a decision produced the same final analysis after conditioning on the other decisions.

```{r}
#| label: fig-specification-curve
#| fig-cap: !expr paste0("Specification curve for the CBT depression multiverse ($S = ", nrow(multi$multi), "$ unique analyses). The upper panel shows pooled Hedges' $g$ estimates, classified by unadjusted and step-down max-$T$ significance. The lower panel shows the analytical decisions defining each analysis; 'Non-discriminating' indicates that multiple levels produced the same final model.")
#| fig-width: 11
#| fig-height: 9

spec_results <- res$summary_table |>
  mutate(
    inference = case_when(
      p.adj <= 0.05 ~ "Significant after adj",
      p <= 0.05 ~ "Significant before adj",
      TRUE ~ "Not significant"
    ),
    inference = factor(inference, levels = c("Not significant", "Significant before adj", "Significant after adj"))
  ) |>
  arrange(estimate) |>
  mutate(specification = row_number())

spec_decisions <- spec_results |>
  transmute(
    specification,
    `Control condition` = recode(control_display, waitlist = "Waitlist", cau = "Care as usual", all = "Unrestricted", non_discriminating = "Non-discriminating"),
    `Risk of bias` = recode(rob_display, low_only = "Low risk only", low_some_concern = "Low risk / some concerns", all = "Unrestricted", non_discriminating = "Non-discriminating"),
    `Treatment format` = recode(format_display, all = "Unrestricted", non_discriminating = "Non-discriminating"),
    `Outcome rater` = recode(rating_display, `self-report` = "Self-reported", clinician = "Clinician-rated", all = "Unrestricted", non_discriminating = "Non-discriminating"),
    `Within-study correlation` = if_else(rho_display == "non_discriminating", "Non-discriminating", paste0("ρ = ", rho_display)),
    `Heterogeneity estimator` = tau_estimator
  ) |>
  pivot_longer(-specification, names_to = "decision", values_to = "level") |>
  mutate(decision = factor(decision, levels = c("Control condition", "Risk of bias", "Treatment format", "Outcome rater", "Within-study correlation", "Heterogeneity estimator")))

p_estimate <- ggplot(spec_results, aes(specification, estimate, color = inference)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_hline(yintercept = median(spec_results$estimate), linetype = "dashed", alpha = 0.6) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("Not significant" = "grey60", "Significant before adj" = "firebrick", "Significant after adj" = "forestgreen")) +
  labs(x = NULL, y = "Pooled Hedges' g", color = NULL) +
  theme_bw(base_size = 14) + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "top")

p_decisions <- ggplot(spec_decisions, aes(specification, level)) +
  geom_point(size = 0.8, alpha = 0.7) +
  facet_grid(decision ~ ., scales = "free_y", space = "free_y", switch = "y") +
  labs(x = "Specifications ordered by pooled effect", y = NULL) +
  theme_bw(base_size = 11) + theme(strip.placement = "outside", strip.background = element_blank(), strip.text.y.left = element_text(angle = 0, face = "bold"), axis.text.y = element_text(size = 8))

p_estimate / p_decisions + plot_layout(heights = c(2.2, 3))
```
